Our Galactic Midplane Is Probably Wrong, and Not by a Little

### 2019-01-03 paper/ms.tex
Remove commented text
Update info on action comp

\subsection{Summary of Action Computation} \label{ssec:action_comp}
We begin by summarizing one way in which actions are computed in the
literature \Gus{cite}. One first assumes an axisymmetric model of the Galaxy.
This model may itself be a poor description of the Galactic potential,

We use the code \texttt{gala} v0.3 to perform orbit integrations and
conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
which uses an action estimator technique presented by
\citet{2014MNRAS.441.3284S}. Another common alternative is \texttt{galpy}
\citep{2015ApJS..216...29B} using either direct orbit integration or the
\Gus{Stackel Fudge} approximation. We use the default \texttt{MWPotential} as
our potential, which is based on the Milky Way potential available in
\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

@@

% {\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
% phase space measurements have been made for an unprecedented number of stars.
% In the face of such large data, it is worth considering the goal of such
% measurements --- a goal which should inform the usage of such data.

% In this work we considered one such question: given the current position in
% phase space of a single star, what can we infer about that star's past phase
% space trajectory? Current methodology involves using a smoothed, analytic form
% for the Galactic potential (with perhaps crude implementations of spiral arms
% or sattelites), fitting such form to some selection of observations, and
% integrating from the currently measured position in phase space. This orbit
% integration is also commonly accompanied with a transformation into
% action-angle space. Other than the integration, every step of that process
% should be viewed with significant scrutiny.

% The more general problem of measuring the orbits of stars in cosmological
% simulations is essentially unexplored in this work and the literature, and we
% delay discussion to future work.

% \subsection{Dynamical Midplane}
% We now seek to define the midplane in a dynamically consistent manner, which
% we refer to as the ``dynamical midplane''. By this, we mean we would like to
% define the midplane in a manner consistent with some sort of dynamical
% constraint. Our method, which we describe now, is just one of many possible
% ways the dynamical midplane can be defined.

% We define the dynamical midplane based on the principle that perfectly
% circular, planar orbits will trace the midplane. Once the circular, planar
% orbits have been identified, the procedure for finding the dynamical midplane
% is relatively simple. From the measurement of each star's orbit, the midplane
% of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
% each orbit also gives a measure of the: dynamical center of the galaxy,
% velocity of the galaxy, and acceleration of the galaxy.

% We pause briefly to consider general difficulties in measuring stellar orbits
% in a cosmological simulation. It may seem reasonable to place the galaxy in an
% inertial reference frame and then to simply measure the positions of each star
% particle over time. However, it is not generally the case that the center of
% mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
% re-run of the simulations presented here\footnote{The re-run contained a
% cosmic ray bug fix which is unimportant here, but the re-run also output the
% acceleration of each particle, which is not the case for the simulations in
% this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
% for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
% vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
% $22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
% results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
% $t^2$, this presents a significant problem for measuring stellar orbits over
% longer timescales.

% Another option is to recenter and reorient the galaxy at each snapshot.
% However, the centering and orientation procedures are not without
% stochasticity themselves. This stochasticity can easily dominate the magnitude
% of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
% $0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
% $8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
% different centering and orientation procedures), but were not able to cleanly
% measure the orbits of disk stars.

% Here we choose to simplify the problem by considering only perfectly circular,
% planar orbits. Since these orbits have no phase-dependent structure, we can
% work on shorter timescales, i.e. for orbits with significant radial or
% vertical action a large fraction of their orbit would need to be sampled in
% order to measure the orbit accurately. Because we know the analytic form of
% these orbits, we can forward model the trajectory of each star particle in the
% cosmological simulation. Working in comoving coordinates, we can write down
% the trajectory of a star particle as
% \beq\label{eq:helix}
% \bm{r}(t) = R(\theta, \phi, \psi) \left[ (A\cos{\omega t}, B\sin{\omega t}, C
% t) + \bm{x_0} \right]\text{,}
% \eeq
% where $R(\theta, \phi, \psi)$ is an Euler angle transformation, $A$, $B$, $C$,
% and $\omega$ are arbitrary constants, and $\bm{x_0}$ is an arbitrary 3-vector.
% This equation is a helix with an arbitrary constant and rotation. It describes
% the

% We minimize the summed squared difference between Equation~\ref{eq:helix} and
% the true trajectory of each star with $R$ within $1\,\kpc$ of $R_{\odot} =
% 8.2\,\kpc$ in each simulation. We consider minimum rms error (i.e. the typical
% error in the Equation~\ref{eq:helix} model in describing the true trajectory).
% This rms error characterizes how well the helix model we consider, which would
% be exactly correct if the overall motion of the galaxy is well-described by
% constant speed motion and the orbit of a particular store is exactly circular
% and planar.

% We plot a histogram of the rms error for each galaxy in
% Figure~\ref{fig:something}. As we estimated before, the total effect of the
% galaxy's acceleration is roughly $90\,\pc$ (\Gus{Check for m12f, m12m}). We
% find that this is small compared to the typical error, even for the stars
% whose trajectories are best modeled by Equation~\ref{eq:helix}.

% There are many reasons for the failure of this model. First, there could be no
% stars on circular orbits. This seems unlikely considering that typical
% distribution functions peak on perfectly circular, planar orbits
% \citep[e.g.][]{2013MNRAS.434..652T}. Second, the trajectory of the galaxy may
% not be well-described by constant-speed motion. This also seems unlikely,
% considering that the bulk acceleration of the stellar disk only results in a
% $\sim90\,\pc$ (\Gus{m12f, m12m}) deviation from constant-speed motion over the
% $22\,\Myr$ time we considered.

% Furthermore, in Figure~\ref{fig:something} we plot the center of mass of the
% galaxy (in comoving coordinates) for the $22\,\Myr$ timeframe described. The
% center of mass at each snapshot is computed using the iterative shrinking
% spheres method described in Section~\ref{ssec:cosmozoom}. The rms error on the
% constant-speed motion is $\sim\text{something}$, which we interpret as
% \Gus{something}.

% One potential explanation is that {\em no star can remain on a circular,
% planar orbit}. The order unity density variations may be strong enough such
% that each star's orbit is perturbed. If there are truly stars in the
% simulation on perfectly circular, planar orbits, then the magnitude of their
% acceleration over the last $22\,\Myr$ must be constant. Furthermore, their
% trajectory in acceleration space must be well-described by a rotation. In the re-runs,

% \section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
% An analytic, axisymmetric potential with a known coordinate system is a far
% too simple description of the Milky Way. We thus turn to cosmological zoom-in
% simulations as approximations to a ``true'' Milky Way potential. We will
% perform a few simulations of small clusters embedded in the potential of a
% frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
% frame from \citet{2018arXiv180610564S}.

% Freezing the potential means that during the course of our simulation the
% cluster will rapidly pass by much of the substructure in the potential. To
% illustrate that this has little effect on our results, we resimulate each
% cluster with the softening lengths of the embedded potential increased by a
% factor of \Gus{something}.

% Once we perform this simulation, we calculate the actions of each star in the
% cluster assuming the best-fit smooth, axisymmetric potential, as we do
% observationally for the Milky Way. We refer to this space as the observational
% action space. We will find that there is significant time evolution in
% observational action space. Attributing some of this evolution to our fiducial
% coordinate system being incorrect, we minimize the time-dispersion in action
% space

% \subsection{Cluster Simulations} \label{ssec:cluster_sim}
% We simulate a small star cluster of $256$ stars in the frozen $\z=0$
% potential of m12i. We use the code \texttt{ph4} within the
% \texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
% 2013A&A...557A..84P} for the cluster integration. We use a softening length
% of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
% the cluster, but this is not important for our purposes. To couple the
% gravitational forces of the frozen snapshot to the cluster stars, we use the
% \texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
% \texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
% $\sim 10$ orbits.

% To compute the forces from the frozen potential, we use the simple tree code
% \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
% commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
% at distances $R>50\,\kpc$ in the fiducial coordinate system.

% We use the initial position and velocity of two star particles in
% m12i. These two star particles are randomly chosen from all star
% particles in the simulation with the following properties (in the fiducial
% coordinate system):
% \begin{enumerate}
%     \item age in the range $(0.25, 0.75)\,\Gyr$,
%     \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
%     \item height (absolute value) less than $0.5\,\kpc$,
%     \item $J_r$ and $J_z$ in the range:
%         \subitem $(10, 20)\,\actunit$ for the first star,
%         \subitem $(40, 60)\,\actunit$ for the second star.
% \end{enumerate}
% The ids we choose are \texttt{23693026} for the first star and
% \texttt{17012804} for the second star. We refrain from calling these ``low''
% or ``high'' action stars since the actions are computed in a presumably
% incorrect coordinate system.

% To compute actions, both in the choosing of each of these star particles and
% in what follows, we use the package
% \texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
% commit: \texttt{0e48336}}. Using only the high-resolution particles, we
% compute the symmetrized potential in two components. First, a multipole
% expansion with $20$ radial grid points and a max order of $2$ for the
% spherical-harmonic expansion is used for the dark matter. Second, a
% cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
% max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
% and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
% maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
% cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
% left as default.

% \subsection{Frame Optimization} \label{ssec:frame_opt}
% Figure~\Gus{something} shows the action computed at each time step in the
% fiducial coordinate system. Clearly there is significant time evolution.

% Even though we change the center, we do not recalculate the center velocity.
% This is largely irrelevant, as the recalculated center velocity differs from
% the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
% package \texttt{gizmo\_analysis} will subtract a slightly different value for
% the Hubble flow from each particle's velocity to convert to a physical
% coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
% also ignore.

>>>>

\subsection{Summary of Action Computation} \label{ssec:action_comp}
We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
We use the code \texttt{gala} v0.3 to perform orbit integrations and
conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
which uses an action estimator technique presented by
\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as
our potential, which is based on the Milky Way potential available in
\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

Other methods for computing actions are used in the literature. For example,
the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
St\"ackel potential to approximate the Galactic potential
\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.

@@
